Simulation of 2D physics of hand drawn objects using OpenCV and Box2D

### 2011-02-07 latex/paper.tex
popisana prva cast listingu a teselacia

Simple image capture is shown in Listing \ref{Capture}.
\begin{lstlisting}[caption=Query image frame from web camera, label=Capture]
self.camera = cv.CaptureFromCAM(-1) %\label{Capture:init})
self.image = cv.QueryFrame(self.camera) %\label{Capture:query})
self.DetectOutline(self.image)
\end{lstlisting}
In line \ref{Capture:init} of listing \ref{Capture} we initialize our web
camera. In variable camera is allocated and initialized object that can query
camera for new image. Then as we see in \ref{Capture:query} we can get the
image from camera and store it in the variable named image. Now when we have
image data stored in the variable, we can process data to find outlines.
\begin{lstlisting}[caption=Outline detection, label=Detect]
def DetectOutline(self, image):
  image_size = cv.GetSize(image)
  grayscale = cv.CreateImage(image_size, 8, 1)	%\label{Detect:grayscale})
  cv.CvtColor(image, grayscale, cv.CV_BGR2GRAY)
  cv.EqualizeHist(grayscale, grayscale)	%\label{Detect:equalize})
  storage = cv.CreateMemStorage(0)
  cv.Threshold(grayscale, grayscale, 50, 255, cv.CV_THRESH_BINARY) %\label{Detect:threshold})
  self.contours = cv.FindContours(grayscale, %\label{Detect:contours})
    cv.CreateMemStorage(),
    cv.CV_RETR_TREE,
    cv.CV_CHAIN_APPROX_SIMPLE)
  if len(self.contours) > 0:
    self.contours = cv.ApproxPoly (self.contours, %\label{Detect:approx})
      storage,
      cv.CV_POLY_APPROX_DP,
      1.5,
      1)
  return self.contours
\end{lstlisting}
In function in Listing \ref{Detect} is shown how to find outlines of objects in
image. We convert image to gray scale as seen on line \ref{Detect:grayscale}.
Then we run histogram equalization (line: \ref{Detect:equalize}).
Equalization makes objects better visible and gives better output for
thresholding (line: \ref{Detect:threshold}) which makes black and white image
prepared for outline detection (line: \ref{Detect:contours}).
%TODO Doplnit obrazok s normalizaciou histogramu
%TODO Doplnit obrazok s thresholdom
After outline detection we have tree of contours stored in the variable
self.contours. These trees are iterable objects sorted from outer to inner
outline connected by property h\_next and v\_next that we will describe in
paragraph about creation of objects from outlines.

@@

doSleep = True %\label{World:sleep})
self.world = box2d.b2World(self.worldAABB, gravity, doSleep)
\end{lstlisting}
First we have to create boundaries of the world. We define them as vectors from
bottom left (line: \ref{World:lower}) to top right (line:
\ref{World:upper}). Objects have to be inside the boundaries, when an object
touch the boundary it gets stuck. Then we define gravity vector (line:
\ref{World:gravity}). The last thing before creation of the world we allow
objects to sleep(line: \ref{World:sleep}). Object that are not moving fall
asleep, then are ignored by the engine. Last line of Listing \ref{World} creates
the world.

World is created and we are ready to create objects from outlines.
\subsection{Objects}
Every object that is simulated in Box2D consists of body and shapes. Our objects
are described by the contour tree. To create objects we need to iterate contour
tree, find contours that belongs to together and create objects for these
contours.
\subsubsection{Contour tree}
Contour tree is object in which are stored points of each contour. Contours are
connected by functions returning reference to other contours with h\_next() and
v\_next(). h\_next() is referencing to deeper contour, and v\_next is
referencing to another object contour. To iterate over all contours we have
created recursive function shown in Listing: \ref{CreateObjects}.
\begin{lstlisting}[label=CreateObjects,caption=Function to iterate through
contour tree]
def CreateObjectsFromCountours(self,cont,h=0,v=0):
  if v>0:	%\label{CreateObjects:depth})
    density = 10.0 %\label{CreateObjects:density10})
  else:
    density = 0 %\label{CreateObjects:density0})
  if len(cont)>8:
    self.CreateObject(cont,h,v)

@@

  if cont.h_next():
    h += 1
    self.CreateObjectsFromCountours(cont.h_next(),h,v)
\end{lstlisting}
We iterate through contour tree. First level is outer contour of the image
(Line: \ref{CreateObjects:depth}), because we do not want the outer contour to
move, we set it as static byt setting density to 0 (Line:
\ref{CreateObjects:density0}). Every other contour is dynamic body with density
set to 10 (Line: \ref{CreateObjects:density10}).

@@

Listing \ref{Objects} shows how to create object from outer contour: 
\begin{lstlisting}[name=Objects,label=Objects,caption=Creation of object from
outer contour]
def CreateObject(self, cont, h,v):
  contM = []
  for point in cont:
    x = point[0]/30.0	%\label{Objects:scaling})
    y = point[1]/30.0
    contM.append((x,y))

@@

  if v==0:
    body = self.world.CreateBody(bd)
    try:
      self.contourBodies.append(body)
    except:
      self.contourBodies = [body] %\label{Objects:CountourBodies})
      body.CreateShape(edgeDef) %\label{Objects:AttachShape})
\end{lstlisting}
Image size is measured in pixels and Box2D units are kilograms, meters,
and seconds (KMS) we should scale images coordinates to fit in 0.1m to
10m. In that scale is performance of Box2D the best. We are doing it by
dividing of value of pixel coordinates by 30.0 (Line: \ref{Objects:scaling})

Then we create a body definition that will represent our contour(Line:
\ref{Objects:CreateBody}) and set up it initial possition in next line.

After that we create shape of body as chain of edges (Line:
\ref{Objects:CreateChain}) and assign the array of vertices to it (Line:
\ref{Objects:AssignVertices}). Edges are special type of shapes that have no
mass, is represented as line betven vertices that collide with other non-edge
objects. Edges are easy to create because they do not have to be concave unlike
polygons.

@@

Listing of the function CreateObject() continous in Listing \ref{Objects}. This
part of function creates dynamic objects inside the outer contour.
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects,caption=Creation of objecs] if v == 1: try:
      body = self.objectBodies[h]
    except:
      body = self.world.CreateBody(bd)
      self.objectBodies[h] = body
    polyline = []
    for (x,y) in cont:
      polyline.append(p2t.Point(x,y))
    cdt = p2t.CDT(polyline)
    triangles = cdt.triangulate()
    for t in triangles:
      x1 = t.a.x/30.0
      y1 = t.a.y/30.0
      x2 = t.b.x/30.0
      y2 = t.b.y/30.0
      x3 = t.c.x/30.0
      y3 = t.c.y/30.0
      if math.hypot(x2-x1,y2-y1)<0.1:
        x2 = x2 + math.copysign(0.1, x2-x1)
        y2 = y2 + math.copysign(0.1, y2-y1)
        print "kratky vektor"
      if math.hypot(x3-x2,y3-y2)<0.1:
        x3 = x3 + math.copysign(0.1, x3-x2)
        y3 = y3 + math.copysign(0.1, y3-y2)
        print "kratky vektor"
      if math.hypot(x1-x3,y1-y3)<0.1:
        x1 = x1 + math.copysign(0.1, x1-x3)
        y1 = y1 + math.copysign(0.1, y1-y3)
        print "kratky vektor"
      poly=box2d.b2PolygonDef()
      poly.setVertices(((x1, y1), (x2, y2), (x3, y3)))
      poly.density = 1.0
      poly.restitution = 0.0
      poly.friction = 0.0
      body.CreateShape(poly)
      body.SetMassFromShapes()
\end{lstlisting}

\subsection{Tesselation}
Box2D supports only collisions between convex objects. That is why we need to
break outlines to convex polygons. There is more ways how to break concave
objects. We chose the 2D constrained Delaunay triangulation algorithm
implemented by poly2tri Python library[\cite{Delauanay}].

>>>>

Simple image capture is shown in Listing \ref{Capture}.
\begin{lstlisting}[caption=Query image frame from web camera, label=Capture]
self.camera = cv.CaptureFromCAM(-1) %\label{Capture:init})
self.image = cv.QueryFrame(self.camera) %\label{Capture:query})
self.DetectOutline(self.image)
\end{lstlisting}
In line \ref{Capture:init} of listing \ref{Capture} we initialize our web
camera. In variable camera is allocated and initialized object that can query
camera for new image. Then as we see in \ref{Capture:query} we can get the
image from camera and store it in the variable named image. Now when we have
image data stored in the variable, we can process data to find outlines.
\begin{lstlisting}[caption=Outline detection, label=Detect]
def DetectOutline(self, image):
  image_size = cv.GetSize(image)
  grayscale = cv.CreateImage(image_size, 8, 1)	%\label{Detect:grayscale})
  cv.CvtColor(image, grayscale, cv.CV_BGR2GRAY)
  cv.EqualizeHist(grayscale, grayscale)	%\label{Detect:equalize})
  storage = cv.CreateMemStorage(0)
  cv.Threshold(grayscale, grayscale, 50, 255, cv.CV_THRESH_BINARY) %\label{Detect:threshold})
  self.contours = cv.FindContours(grayscale, %\label{Detect:contours})
    cv.CreateMemStorage(),
    cv.CV_RETR_TREE,
    cv.CV_CHAIN_APPROX_SIMPLE)
  if len(self.contours) > 0:
    self.contours = cv.ApproxPoly (self.contours, %\label{Detect:approx})
      storage,
      cv.CV_POLY_APPROX_DP,
      1.5,
      1)
  return self.contours
\end{lstlisting}
In function in Listing \ref{Detect} is shown how to find outlines of objects in
image. We convert image to gray scale as seen on line \ref{Detect:grayscale}.
Then we run histogram equalization (line:~\ref{Detect:equalize}).
Equalization makes objects better visible and gives better output for
thresholding (line:~\ref{Detect:threshold}) which makes black and white image
prepared for outline detection (line:~\ref{Detect:contours}).
%TODO Doplnit obrazok s normalizaciou histogramu
%TODO Doplnit obrazok s thresholdom
After outline detection we have tree of contours stored in the variable
self.contours. These trees are iterable objects sorted from outer to inner
outline connected by property h\_next and v\_next that we will describe in
paragraph about creation of objects from outlines.

@@

doSleep = True %\label{World:sleep})
self.world = box2d.b2World(self.worldAABB, gravity, doSleep)
\end{lstlisting}
First we have to create boundaries of the world. We define them as vectors from
bottom left (line:~\ref{World:lower}) to top right (line:~
\ref{World:upper}). Objects have to be inside the boundaries, when an object
touch the boundary it gets stuck. Then we define gravity vector (line:~
\ref{World:gravity}). The last thing before creation of the world we allow
objects to sleep(line:~\ref{World:sleep}). Object that are not moving fall
asleep, then are ignored by the engine. Last line of Listing \ref{World} creates
the world.

World is created and we are ready to create objects from outlines.
\subsection{Objects}
Every object that is simulated in Box2D consists of body and shapes. Our objects
are described by the contour tree. To create objects we need to iterate contour
tree, find contours that belongs to together and create objects for these
contours.
\subsubsection{Contour tree}
Contour tree is object in which are stored points of each contour. Contours are
connected by functions returning reference to other contours with h\_next() and
v\_next(). h\_next() is referencing to deeper contour, and v\_next is
referencing to another object contour. To iterate over all contours we have
created recursive function shown in Listing: \ref{CreateObjects}.
\begin{lstlisting}[label=CreateObjects,caption=Function to iterate through
contour tree]
def CreateObjectsFromCountours(self, cont, h=0, v=0):
  if v>0:	%\label{CreateObjects:depth})
    density = 10.0 %\label{CreateObjects:density10})
  else:
    density = 0 %\label{CreateObjects:density0})
  if len(cont)>8:
    self.CreateObject(cont,h,v)

@@

  if cont.h_next():
    h += 1
    self.CreateObjectsFromCountours(cont.h_next(),h,v)
\end{lstlisting}
We iterate through contour tree. First level is outer contour of the image
(line:~\ref{CreateObjects:depth}), because we do not want the outer contour to
move, we set it as static byt setting density to 0 (line:~
\ref{CreateObjects:density0}). Every other contour is dynamic body with density
set to 10 (line:~\ref{CreateObjects:density10}).

@@

Listings \ref{Objects1},\ref{Objects2},\ref{Objects3} shows how to create object
from outer contour:
\begin{lstlisting}[name=Objects,label=Objects1,caption=Creation of object from
outer contour]
def CreateObject(self, cont, h,v):
  contM = []
  for point in cont:
    x = point[0]/30.0	%\label{Objects:scaling})
    y = point[1]/30.0
    contM.append((x,y))

@@

  if v==0:
    body = self.world.CreateBody(bd)
    try:
      self.contourBodies.append(body)
    except:
      self.contourBodies = [body] %\label{Objects:CountourBodies})
      body.CreateShape(edgeDef) %\label{Objects:AttachShape})
\end{lstlisting}
Image size is measured in pixels and Box2D units are kilograms, meters,
and seconds (KMS) we should scale images coordinates to fit in 0.1m to
10m. In that scale is performance of Box2D the best. We are doing it by
dividing of value of pixel coordinates by 30.0 (line:~\ref{Objects:scaling})

Then we create a body definition that will represent our contour(line:~
\ref{Objects:CreateBody}) and set up it initial possition in next line.

After that we create shape of body as chain of edges (line:~
\ref{Objects:CreateChain}) and assign the array of vertices to it (line:~
\ref{Objects:AssignVertices}). Edges are special type of shapes that have no
mass, is represented as line betven vertices that collide with other non-edge
objects. Edges are easy to create because they do not have to be concave unlike
polygons.

@@

Listing of the function CreateObject() continous in Listing \ref{Objects2}. This
part of function creates dynamic objects inside the outer contour. In this part
we prepare list for bodies of objects, so we can modify objects that are already
created or objects that we want append new shapes.
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects2,caption=Creation of objects] if v == 1:
    try:
      body = self.objectBodies[h]
    except:
      body = self.world.CreateBody(bd)
      self.objectBodies[h] = body
\end{lstlisting}
\subsection{Tesselation}
Box2D supports only collisions between convex objects and contours of objects
captured by camera are not convex we have to break outlines to convex polygons.
There is more ways how to break concave objects. We chose the 2D constrained
Delaunay triangulation algorithm implemented by poly2tri Python
library[\cite{Delauanay}]. Function CreateObject() continous in Listing~
\ref{Objects3}
\begin{lstlisting}[name=Objects,firstnumber=auto,label=Objects3,caption=Creation
 of objects]
    polyline = []
    for (x,y) in cont:
      polyline.append(p2t.Point(x,y)) %\label{Objects:polyline})
    cdt = p2t.CDT(polyline) %\label{Objects:p2t})
    triangles = cdt.triangulate()
    for t in triangles:
      x1 = t.a.x/30.0
      y1 = t.a.y/30.0
      x2 = t.b.x/30.0
      y2 = t.b.y/30.0
      x3 = t.c.x/30.0
      y3 = t.c.y/30.0
      if math.hypot(x2-x1,y2-y1)<0.1: %\label{Objects:minSize1})
        x2 = x2 + math.copysign(0.1, x2-x1)
        y2 = y2 + math.copysign(0.1, y2-y1)
      if math.hypot(x3-x2,y3-y2)<0.1:
        x3 = x3 + math.copysign(0.1, x3-x2)
        y3 = y3 + math.copysign(0.1, y3-y2)
      if math.hypot(x1-x3,y1-y3)<0.1:
        x1 = x1 + math.copysign(0.1, x1-x3)
        y1 = y1 + math.copysign(0.1, y1-y3) %\label{Objects:minSize2})
      poly=box2d.b2PolygonDef()  %\label{Objects:polygondef})
      poly.setVertices(((x1, y1), (x2, y2), (x3, y3)))
      poly.density = 1.0
      poly.restitution = 0.0
      poly.friction = 0.0
      body.CreateShape(poly)  %\label{Objects:createshape})
      body.SetMassFromShapes() %\label{Objects:setMass})
\end{lstlisting}
Objects:polyline
Objects:p2t
Objects:minSize1
Objects:minSize2
Objects:polygondef
Objects:createshape
Objects:setMass
